Citation Analysis of Computer Science Publications

### 2019-11-24 Scientometrics/comp.tex
updated

\title{Citation Analysis of Computer Science%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Relationship With Other Fields}

@@

\begin{abstract}
Computer science as a field has experienced dramatic growth and diversification over the last twenty years. Towards an improved understanding of the structure of the field, we analyze a cohort of the computer science literature using a historiographic approach. For further insight on the features of this cohort and the flow of information within its components, we construct article level clusters whose elements are linked by citations or co-citations, and reconcile them with subject categories in the All Science Journal Classification (Scopus). [abstract is feeble and needs to be rewritten]

@@

In extending the work of Salton and Bergmark to the 21st century, we were also interested in connections between computer science and other fields. As a source of computer science literature, we used DBLP, a reference bibliography for computer science~\cite{dblp_ref}. The DBLP bibliography covers publications from computer science and includes publications from hybrid fields in cases they are considered pertinent to computer science research. The underlying assumption is that records in DBLP are greatly enriched for computer science even while recognizing that coverage is incomplete and possibly biased. Cross-matching DBLP publications to Scopus records using whole text searches applied to titles, enabled harvesting citation data as well as publications from other disciplines. This cross-matching also allows the use of both journal and article based classifications when clustering documents. Using both direct citations and co-citations as the basis for clustering, we used articles from from journals and conference proceedings for the period 1996-2015. We reconciled these clusters with the the All Science Journal Classification (ASJC) developed and maintained by Scopus as well as to the CCS classification through a combination of automated and manual procedures. We describe below the high-level interactions of computer science within the Physical Sciences as well as with the Social Sciences and Life Sciences.

@@

\subsection{Data}
 A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the DBLP dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning vectors containing these lexemes. The vectors are then matched with corresponding text search vectors extracted from DBLP document titles. A rank is returned based on the frequency of matching lexemes. Only records in DBLP with DOIs (4,291,130) were considered and only matches with a normalized rank of greater than 0.99 were accepted, roughly equivalent to `at least 11 matching lexemes independent of order and vector length'. This procedure resulted in a dataset of 634,100 DBLP publications with Scopus identifiers where 49.6\% were labeled as article and 48.8\% as conference proceeding (Fig~\ref{fig:ar_cp_annotation}). Cited references were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations within the dataset. This dataset is referred to as \emph{comp}.\par

 \begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.5]{ar_cp_ratio.pdf}
% figure caption is below the figure
\caption{Summary of DBLP data cross-matched with Scopus. 634,100 publications from DBLP were cross-matched with Scopus and then matched to the 27 major subject areas in the ASJC (Scopus) classifications. The largest number of publications are contributed by Computer Science; Mathematics; and Engineering and then by Social Sciences; Physics and Astronomy; and Biochemistry, Genetics, and Molecular Biology. Publications were further annotated with respect to being either articles or conference proceedings and ordered on the abscissa by the ratio of conference proceedings to articles within a major subject area (cp\_count/ar\_count). For this dataset, the major subject areas of Energy; Computer Science; Medicine; Mathematics; and Engineering had the highest ratio of conference proceedings to articles while Immunology and Microbiology; Chemistry; General (multidisciplinary); Pharmacology, Toxicology, and Pharmaceutics; and Agricultural and Biological Sciences had the lowest ratio. }
\label{fig:ar_cp_annotation}       % Give a unique label
\end{figure}

@@

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments.
The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus. We varied the number of clusters to be formed between 10 and 50 clusters. At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}. Consequently, we generated 20 clusters using Graclus, which were labeled 0-19 (Fig~\ref{fig:heatmap}).

@@

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.5]{scopus_dblp_graclus2.pdf}
% figure caption is below the figure
\caption{caption will be inserted I just haven't been able to overcome the threshold to avoid boring tasks}
\label{fig:heatmap}       % Give a unique label
\end{figure}
For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography published in the years 1996-2015, that had a digital object identifier (DOI), and could also be matched to article identifiers in the Scopus bibliography. The process of selection and matching resulted in a dataset of 634,100 publications (Materials and Methods). 
INSERT sentence about why one should cluster this dataset. Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with an argument being made for direct citation as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using a spectral clustering technique, Graclus, that has previously been applied to citation data~\cite{graclus_2007,subelj_clustering_2016}. We constructed article level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to disciplinary classification by journal, the Scopus ASJC classification. We identified ASJC categories that accounted for at least 10\% of the publications in each cluster and observed 31 of the 334 minor subject areas in the Scopus ASJC classification, which consists of 5 subject areas, 27 major subject areas, and 334 minor subject areas (Fig~\ref{fig:heatmap}).

@@

% For one-column wide figures use
\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:1}       % Give a unique label
\end{figure}
%
% For two-column wide figures use
\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[width=0.75\textwidth]{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:2}       % Give a unique label
\end{figure*}
%
% For tables use
\begin{table}
% table caption is above the table
\caption{Please write your table caption here}
\label{tab:1}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lll}
\hline\noalign{\smallskip}
first & second & third  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
number & number & number \\
number & number & number \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

>>>>

\title{Citation Analysis of Computer Science Publications%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Relationship To Other Fields}

@@

\begin{abstract}
Computer science as a field has experienced dramatic growth and diversification over the last twenty years. Towards an improved understanding of the structure of the field, we analyze a cohort of the computer science literature.  For insight on the features of this cohort and the flow of information within its components, we construct article level clusters whose elements are linked by citations or co-citations, and reconcile them with subject categories in the All Science Journal Classification (Scopus). [abstract is feeble and needs to be rewritten]

@@

Thus, we sought to extend the work of Salton and Bergmark to reflect a more current reality while also considering connections between computer science and other fields. As a source of computer science literature, we used DBLP, a reference bibliography for computer science~\cite{dblp_ref}. The DBLP bibliography covers publications from computer science and includes publications from hybrid fields, where they are considered pertinent to computer science research. The underlying assumption is that records in DBLP are greatly enriched for computer science even while recognizing that coverage is incomplete and possibly biased. Cross-matching DBLP publications to Scopus records using whole text searches applied to titles, enabled harvesting citation data as well as publications from other disciplines. This cross-matching also allows the use of both journal and article based classifications when clustering documents. Using both direct citations and co-citations as the basis for clustering, we used articles from from journals and conference proceedings for the period 1996-2015. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus as well as to the CCS classification through a combination of automated and manual procedures. We describe below the high-level interactions of computer science within the Physical Sciences as well as with the Social Sciences and Life Sciences.

@@

\subsection{Data}
 A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the DBLP dataset, were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceeding (Fig~\ref{fig:ar_cp_annotation}). References cited by these publications were then extracted from Scopus (7,129,006), resulting in a total of 8,000,411 nodes (publications and references) and  44,296,381 edges that represent citations within the dataset. This dataset is referred to as \emph{comp}.\par

 \begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.5]{ar_cp_ratio.pdf}
% figure caption is below the figure
\caption{Summary of DBLP data cross-matched with Scopus. 2,685,356 publications from DBLP were cross-matched with Scopus and then grouped by the 27 major subject areas in the ASJC (Scopus) classification. The largest number of publications are contributed by Computer Science; Mathematics; and Engineering and then by Social Sciences; Physics and Astronomy; and Biochemistry, Genetics, and Molecular Biology. Publications were further annotated with respect to being either articles \emph{(ar)} or conference proceedings \emph{(cp)}. For this dataset, the major subject areas of Computer Science; Engineering; and Mathematics contributed the most publications while Pharmacology, Toxicology, and Pharmaceutics; Nursing, Veterinary and Dentistry contributed the least.}
\label{fig:ar_cp_annotation}       % Give a unique label
\end{figure}

@@

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments.
The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus. We varied the number of clusters to be formed between 10 and 50 clusters. At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}. Consequently, we generated 20 clusters using Graclus, which were labeled 0-19 (Fig~\ref{fig:heatmap})

\begin{table}
% table caption is above the table
\caption{\emph{2685021 is because of inner joins on ASJC Explain conductance calculations~\cite{shun_parallel_2016}}}
\label{tab:1}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lllll}
\hline\noalign{\smallskip}
Cluster & Publications & Conductance & ASJC Labels & Unique Labels\\
\noalign{\smallskip}\hline\noalign{\smallskip}
0 & 111,294 & 0.12 & 265,664 & 142 \\ 
1 & 117,057 & 0.14 & 246,960 & 166 \\ 
2 & 116,251 & 0.11 & 280,602 & 165 \\ 
3 & 353,366 & 0.15 & 881,693 & 200 \\ 
4 & 145,081 & 0.09 & 349,020 & 154 \\ 
5 & 92,097 & 0.17 & 248,168 & 186 \\ 
6 & 71,865 & 0.15 & 199,681 & 163 \\ 
7 & 179,927 & 0.18 & 465,181 & 202 \\ 
8 & 302,656 & 0.18 & 760,117 & 214 \\ 
9 & 69,520 & 0.19 & 197,031 & 174 \\ 
10 & 42,462 & 0.10 & 102,838 & 141 \\ 
11 & 448,030 & 0.25 & 1,229,061 & 224 \\ 
12 & 70,738 & 0.21 & 216,546 & 179 \\ 
13 & 105,232 & 0.17 & 289,318 & 187 \\ 
14 & 199,176 & 0.15 & 551,657 & 208 \\ 
15 & 64,384 & 0.17 & 195,679 & 167 \\ 
16 & 89,340 & 0.11 & 240,157 & 158 \\ 
17 & 50,817 & 0.14 & 181,531 & 179 \\ 
18 & 43,113 & 0.20 & 108,518 & 177 \\ 
19 & 12,615 & 0.80 & 36,583 & 229 \\ \noalign{\smallskip}\hline
\end{tabular}
\end{table}

@@

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.5]{scopus_dblp_graclus3.pdf}
% figure caption is below the figure
\caption{caption will be inserted I just haven't been able to overcome the threshold to avoid boring tasks}
\label{fig:heatmap}       % Give a unique label
\end{figure}
For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography published in the years 1996-2015, that had a digital object identifier (DOI), and could also be matched to article identifiers in the Scopus bibliography. The process of selection and matching resulted in a dataset of 634,100 publications (Materials and Methods). 
INSERT sentence about why one should cluster this dataset. Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with an argument being made for direct citation as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using a spectral clustering technique, Graclus, that has previously been applied to citation data~\cite{graclus_2007,subelj_clustering_2016}. We constructed article level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to disciplinary classification by journal, the Scopus ASJC classification. We identified ASJC categories that accounted for at least 10\% of the publications in each cluster and observed 31 of the 334 minor subject areas in the Scopus ASJC classification, which consists of 5 subject areas, 27 major subject areas, and 334 minor subject areas (Fig~\ref{fig:heatmap}).

@@

% For one-column wide figures use
\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:1}       % Give a unique label
\end{figure}
%
% For two-column wide figures use
\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[width=0.75\textwidth]{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:2}       % Give a unique label
\end{figure*}
%
% For tables use
